---
title: "beautiful graphics ggplot2"
author: "C.J. Liu"
date: "11/17/2016"
output: 
    html_document:
        toc: true
        toc_float:
            collapsed: false
            smooth_scroll: false
        depth: 3
        theme: united
        number_sections: true
        highlight: tango
---

# Plot One Variable - X: Continuous or Discrete

For **one continuous variable** = **Numeric**:

* geom_area()
* geom_density()
* geom_histogram()
* geom_freqpoly()
* geom_dotplot()
* stat_ecdf()
* stat_qq()

For **one discrete varaible** = **Factor**:

* geom_bar()

```{r load library}
library(ggplot2)
library(dplyr)
```


## Area Plots
**alpha, color, fill, linetype, size**
```{r}
set.seed(1234)
wdata = as_data_frame(data.frame(sex = factor(rep(c("F", "M"), each=200)), weight = c(rnorm(200,55),rnorm(200,58))))
wdata

mu <- wdata %>% group_by(sex) %>% summarize(grp.mean = mean(weight))
mu

a <- ggplot(wdata, aes(x = weight))

a + geom_area(stat = "bin", color = "black", fill = "#00AFBB")
#a + geom_area() will not get right result, object 'y' not found. Use stat to specify the count as y 
#Note that, by default y axis corresponds to the count of weight values. If you want to change the plot in order to have the density on y axis, the R code would be as follow.
a + geom_area(aes(y = ..density..), stat = "bin")

data("diamonds")
diamonds <- as_data_frame(diamonds)
diamonds
p <- ggplot(diamonds, aes(x = price, fill = cut))
# Bar plot
p + geom_bar(stat = "bin")
# Area plot
p + geom_area(stat = "bin")
```

## Density Plots
**alpha, color, fill, linetype, size**

* scale_color_manual(), scale_fill_manual()
* scale_color_brewer(), scale_fill_brewer() RColor-Brewer
* scale_color_grey(), scale_fill_grey()


```{r}
# Basic plots
a + geom_density()

# Add color and mean xintercept and median xintercept
a + geom_density(color = "black", fill = "gray") + geom_vline(aes(xintercept = mean(weight)), color = "#FC4E08", linetype = "dashed", size = 1) + geom_vline(aes(xintercept = median(weight)), color = "blue", linetype = 4, size = 1)

# Change color by group
a + geom_density(aes(fill = sex), alpha = 0.4) 

# Add mean lines and color by sex
a + geom_density(aes(fill = sex), alpha = 0.4) + geom_vline(data = mu, aes(xintercept = grp.mean, color = sex), linetype = "dashed")

# Change manually 
# change line manually
a2 <- a + geom_density(aes(color = sex)) + geom_vline(data = mu, aes(xintercept = grp.mean, color = sex), linetype = "dashed") + theme_minimal()

a2 + scale_color_manual(values = c("#999999", "#E69F00"))

a2 + scale_color_brewer(palette = "Paired")

a2 + scale_color_grey()

# change fill manually
a3 <- a + geom_density(aes(fill = sex), alpha = 0.4) + theme_minimal()

a3 + scale_fill_manual(values = c("#999999", "#E69F00"))

a3 + scale_fill_brewer(palette = "Dark2")

a3 + scale_fill_grey()
```

## Histogram Plots
**identity(position\_identity()), stack(position\_stack()), dodge(position\_dodge()); Default values is "stack"**

**alpha, color, fill, linetype, size**
```{r}
# Basic plot
a + geom_histogram()

a + geom_histogram(bins = 50)

#Note that by default, stat_bin uses 30 bins - this might not be good default. You can change the number of bins (e.g.: bins = 50 or the bin width e.g.: binwidth = 0.5.
a + geom_histogram(bins = 50, color = "black", fill = "grey") + geom_vline(aes(xintercept = mean(weight)), color = "#FC4E07", linetype = "dashed", size = 1) + theme_minimal()

a + geom_histogram(aes(y = ..density..), bins = 50)

# Change color by sex
a + geom_histogram(aes(color = sex), fill = "white", bins = 50) + theme_minimal()

# Position adjustment "identity"(overlaid)
a + geom_histogram(aes(color = sex), fill = "white", bins = 50, alpha = 0.6, position = "identity")

# Position adjustment "dodge" (Interleaved)
# Add mean lines and color by sex
a + geom_histogram(aes(color = sex), fill = "white", alpha = 0.6, position = "dodge", bins = 50) + geom_vline(aes(xintercept = mean(weight)), linetype = "dashed")

# Change fill, color manually
# Change outline color manually
a + geom_histogram(aes(color = sex), fill = "white", alpha = 0.4, position = "identity", bins = 50) + scale_color_manual(values = c("#00AFBB","#E7B800"))

# Change fill and outline color manually
# a + geom_histogram(aes(color = sex), fill = "white", alpha =0.4, position = "identity", bins = 50) + scale_fill_manual(values = c("#00AFBB", "#E7B800")) + scale_color_manual(values = c("#00AFBB", "#E7B800")) 
# wrong command, I have to assign fill first by group

a + geom_histogram(aes(color = sex, fill = sex), alpha =0.4, position = "identity", bins = 50) + scale_fill_manual(values = c("#00AFBB", "#E7B800")) + scale_color_manual(values = c("#00AFBB", "#E7B800")) 


## Combine Histogram and Density Plots

# Plot histogram with density values on y-axis(instead of count values).
# Add density plot with transparent density plot

# Histogram with density plot
a + geom_histogram(aes(y = ..density..),color = "black", fill = "white") + geom_density(alpha = 0.2, fill = "#FF6666") + theme_minimal()

# Color by groups
a + geom_histogram(aes(y = ..density.., color = sex, fill = sex),  alpha = 0.4, position = "identity") + geom_density(aes(color = sex), size =1)


```

## Frequency Polygon
Very close to histogram plots

* Histogram use bars
* Frequency polygons use lines.

**alpha, color, linetype, size**
```{r}
# Basic plot
a + geom_freqpoly(bins = 30) + theme_minimal()

# Change color and linetype by sex
# Use custom color palettes
a + geom_freqpoly(aes(color = sex, linetype = sex), bins = 30 ) +  scale_color_manual(values = c("#999999", "#E69F00"))+theme_minimal()

# y density
a + geom_freqpoly(aes(y = ..density.., color = sex, linetype = sex), bins = 30 ) +  scale_color_manual(values = c("#999999", "#E69F00"))+theme_minimal()
```

## Dot Plots for One Variable
Not suitable for one variable, it's ugly.
```{r}
a + geom_dotplot(aes(fill = sex))
```

## ECDF Plots
Empirical Cumulative Density Function

**alpha, color, linetype, size**

```{r}
a + stat_ecdf(geom = "point")

a + stat_ecdf(geom = "step")
```

## QQ Plots
Quantile - Quantie plots to chech whether a given data follows normal distribution.

**alpha, color, shape, size**
```{r}
data(mtcars)
mtcars <- as_data_frame(mtcars)
mtcars
mtcars <- mutate(mtcars, cyl = as.factor(cyl))
mtcars

p <- ggplot(mtcars, aes(sample = mpg))

# Basic plot
p + stat_qq()

# Change point shapes by groups
# Use custom color palettes
p + stat_qq(aes(shape = cyl, color = cyl)) + scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))
```

## Bar Plots of Counts
For one discrete variable

**alpha, color, fill, linetype, size**

```{r}
data(mpg)
mpg <- as_data_frame(mpg)
mpg

ggplot(mpg, aes(fl)) + geom_bar(fill = "steelblue") + theme_minimal()
```


# Plot Two Variables -X & Y: Both Continuous or Discrete

## Scatter plots: Continuous X and Y

* geom_pint()
* geom_smooth()
* geom_quantile()
* geom_rug()
* geom_jitter()
* geom_text()

geom_point  
**alpha, color, fill, shape, size**
```{r}
# Data format
mtcars
b <- ggplot(mtcars, aes(x = wt, y= mpg))
# x weight
# y miles/gallon
#Basic scatter plots
b + geom_point(color = "#00AFBB")

# Change the point size, and shape
b + geom_point(color = "#00AFBB", size = 2, shape = 23)

# Control point size by continuous variable values
# qsec 1/4 mile time
b + geom_point(aes(size = qsec), color = "#00AFBB")

# Label text
b + geom_point() + geom_text(label = rownames(mtcars), nudge_y = 0.8)

# Change shape, color, size automatically
# Change point shape by the level of cyl
b + geom_point(aes(shape = cyl))

# Change point shape and colors
b + geom_point(aes(color = cyl, shape = cyl))

# Change shape, color, size manually
# Change the point sizes manually
b + geom_point(aes(color = cyl, shape = cyl, size = cyl)) + scale_size_manual(values = c(2,3,4))

# Change the point shapes and colors manually
b + geom_point(aes(color = cyl, shape = cyl)) + scale_shape_manual(values = c(3,16,17)) + scale_color_manual(values = c('#999999','#E69F00', '#56B4E9'))

# Use brewer color palettes
b + geom_point(aes(color = cyl, shape = cyl)) + scale_color_brewer(palette = "Dark2") + theme_minimal()

# Use grey scale
b + geom_point(aes(color = cyl, shape = cyl)) + scale_color_grey() + theme_minimal()

#####################################################
##Add regression line or smoothed conditional mean###
#####################################################
#geom_smooth(), geom_abline()
#alpha, color, fill, shape, linetype, size
#geom_smooth(method = "auto")
#method:loess->local regression, lm-> linear regression

# Add regression line
b + geom_point() + geom_smooth(method = lm)

# Point + regression line
# Remove the confidence interval
b + geom_point() + geom_smooth(method = lm, se = FALSE)

# loess method, local regression fitting
b + geom_point() + geom_smooth()

# Change the color and shape by groups 吧
b + geom_point(aes(color = cyl, shape = cyl)) + geom_smooth(aes(color = cyl, fill = cyl), method = lm)

# Remove confidence intervals
# Extend the regression lines: fullrage
b + geom_point(aes(color = cyl, shape = cyl)) + geom_smooth(aes(color = cyl), method = lm, se = FALSE, fullrange = TRUE)

# Add marginal rugs to a scatter plot
#geom_rug(sides = "bl")
# sides: a string, "trbl", top, right, bottom, left.
# Add marginal rugs
b + geom_point() + geom_rug()

# Change the color by group
b + geom_point(aes(color = cyl)) + geom_rug(aes(color = cyl))

# Add marginal rugs using faithful data
data(faithful)
faithful <- as_data_frame(faithful)
faithful
ggplot(faithful, aes(x = eruptions, y = waiting)) + geom_point() + geom_rug()

# Jitter points to reduce overplotting
# geom_jitter(), position_jitter()
#alpha, color, fill, shape, size

# Use mpg data
p <- ggplot(mpg, aes(displ, hwy))

# Default sactter plot
p + geom_point()

# Use jitter to reduce overplotting
p + geom_jitter(position = position_jitter(width = 0.5, height = 0.5))
select(mpg, displ, hwy) %>% arrange(-hwy) %>% filter(displ == 1.9)


##
#Text annotation
#geom_text()
#label, alpha, angle, color, family, fontface, hjust, lineheight, size, vjust

b + geom_text(aes(label = rownames(mtcars)), size = 3)

```

## Continuous bivariate distribution

* geom_bin2d()
* geom_hex()
* geom\_density\_2d()


```{r}
c <- ggplot(diamonds, aes(carat, price))
# Add heatmap of 2d bin counts
# geom_bin2d produce a scatter plot with rectangular bins.
# stat_bin_2d(), stat_summary_2d()
# max, xmin, ymax, ymin, alpha, color, fill, linetype, size
c + geom_bin2d()

# Change the number of bins
c + geom_bin2d(bins = 15)

# Specify the width of bins
c + geom_bin2d(binwidth = c(1,1000))
c + stat_bin_2d()
c + stat_summary_2d(aes(z = depth))

# Add hexagon bining
#geom_hex()
# stat_bin_hex(), stat_summary_hex()
# alpha, color, fill, size
require(hexbin)
c + geom_hex()

# Change the number of bins
c + geom_hex(bins = 10)
c + stat_bin_hex()
c + stat_summary_hex(aes(z = depth))

# 2D density estimation
# geom_density_2d()
# stat_density_2d()
# alpha, color, linetype, size

# Scatter plot
sp <- ggplot(faithful, aes(x = eruptions, y = waiting))
select(faithful, eruptions, waiting)
# Default plot
sp + geom_density_2d(color = "#E7B800")
# Add points
sp + geom_point(color = "#00AFBB") + geom_density_2d(color = "#E7B800")
# Use stat_density_2d with geom = "polygon"
sp + geom_point() + stat_density_2d(aes(fill = ..level..), geom = "polygon")
# Change the gradient color
sp + geom_point() + stat_density_2d(aes(fill = ..level..), geom = "polygon") + scale_fill_gradient(low = "#00AFBB", high = "#FC3E07")
# Gradient
```

## Two variables: Discrete X, Discrete Y
geom_jitter  
alpha, color, fill, shape, size  
```{r}
ggplot(diamonds, aes(cut, color)) + geom_jitter(aes(color = cut), size = 0.5)
select(diamonds, cut, color)
```


# Plot Two Variables - X & Y: Discrete X and Continuous Y

* geom_boxplot()
* geom_violin()
* geom_dotplot()
* geom_jitter()
* geom_line()
* geom_bar()

```{r}
data("ToothGrowth")
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
ToothGrowth <- as_data_frame(ToothGrowth)
ToothGrowth
e <- ggplot(ToothGrowth, aes(x = dose, y = len))
```

## Box Plots

**alpha, color, linetype, shape, size, fill**
```{r}
# Basic box plot
e + geom_boxplot()
# Rotate the box plot
e + geom_boxplot() + coord_flip()
# Notched box plot
e + geom_boxplot(notch = TRUE)
# Box plot with mean points
e + geom_boxplot() + stat_summary(fun.y = mean, geom = "point", shape = 18, size = 4, color = "blue")
```


